Choosing TTC Subway Service Provides the Most Time Stability during Weekday Rush-Hours

### 2024-09-24 outputs/paper/paper.qmd
added to data section and reference, re-rendered paper

source of data (brief) structure of data & important attributes

why are these attributes kept for analysis (measurement)

```{r}
#| label: tbl-bus-delay
#| tbl-cap: Sample of Cleaned TTC Bus Delay Dataset
#| echo: false

@@

source of data (brief) structure of data & important attributes

why are these attributes kept for analysis (measurement)

```{r}
#| label: tbl-streetcar-delay
#| tbl-cap: Sample of Cleaned TTC Streetcar Delay Dataset
#| echo: false

@@

source of data (brief)

@@

purpose of extra dataset how it is related to subway_delay

```{r}
#| label: tbl-subway-code
#| tbl-cap: Sample of TTC Subway Delay Code
#| echo: false
#| message: false

@@

### Delay Data

structure of data & important attributes

why are these attributes kept for analysis (measurement) mention "code" column

```{r}
#| label: tbl-subway-delay
#| tbl-cap: Sample of Cleaned TTC Subway Delay Dataset
#| echo: false
#| message: false

@@

```{r}
#| label: fig-total-delay
#| fig-cap: Visualization to show the TTC service that has the highest number of delays in year 2023
#| echo: false
#| message: false
#construct appropriate table for visualization
service_type <- c("bus", "streetcar", "subway")
num_delay <- c(nrow(bus_delay), nrow(streetcar_delay), nrow(subway_delay))

@@

```{r}
#| label: fig-weekday
#| fig-cap: Visualization to show the total delayed time during weekdays in 2023
#| echo: false
#| message: false

@@

```{r}
#| label: tbl-reason
#| tbl-cap: Top 5 Most Popular Reason for Delay
#| echo: false
#| message: false

@@

```{r}
#| label: tbl-lines
#| tbl-cap: Most delayed Lines for each TTC Service
#| echo: false
#| message: false

>>>>

TTC bus delay data are collected and published by the Toronto Transit Commission [@Bus], with latest update in September of 2024. A total of 7 attributes of the data are kept for further evaluation: *date of recorded delay, bus route, scheduled time of arrival, scheduled day of the week (ex. Sunday, Monday, etc.), station, cause of delay,* and *total minutes delayed*. Additional data attributes, such as *vehicle number*, were excluded from the final dataset as they are less relevant to the focus of our analysis—weekday rush-hour delay patterns.

Please see @tbl-bus for an overview of the data.

```{r}
#| label: tbl-bus
#| tbl-cap: Sample of Cleaned TTC Bus Delay Dataset
#| echo: false

@@

Streetcar delay data are also collected and published by the Toronto Transit Commission [@Bus]. Latest update of this dataset occurred in September of 2024. A total of 7 attributes of the data are kept for further evaluation: *date of recorded delay, streetcar line, scheduled time of arrival, scheduled day of the week* (ex. Sunday, Monday, etc.)*, station, cause of delay,* and *total minutes delayed*. Additional data attributes, such as *time gap between current and next vehicle*, were excluded from the final dataset as they are less relevant to the focus of our analysis—weekday rush-hour delay patterns.

@tbl-streetcar contains an overview of streetcar delay data.

```{r}
#| label: tbl-streetcar
#| tbl-cap: Sample of Cleaned TTC Streetcar Delay Dataset
#| echo: false

@@

Publisher of data, update cycle (last updated date)
Streetcar delay data are also collected and published by the Toronto Transit Commission [@Bus]. Latest update of this dataset occurred in September of 2024.

@tbl-subway contains an overview of subway delay data.

### Delay Data {#sec-subway-delay}

A total of 7 attributes of the data are kept from the 2023 subway delay data file for further evaluation: *date of recorded delay, scheduled time of arrival, scheduled day of the week* (ex. Sunday, Monday, etc.)*, station, cause of delay, total minutes delayed,* and *subway line*. One potentially confusing aspect of the subway data is the *cause of delay* attribute, which is represented as codes in the dataset.

An additional dataset containing the code descriptions -- subway delay code -- was extracted and cleaned for reference. Please see @tbl-code for a sample of subway incident codes.

```{r}
#| label: tbl-subway
#| tbl-cap: Sample of Cleaned TTC Subway Delay Dataset
#| echo: false
#| message: false

@@

The table below contains a cleaned dataset used to reference TTC subway delay causes. This delay code dataset contains two key attributes: *code*, and *code description*. Delay code data is used throughout this paper as an reference for the subway delay data discussed in the @sec-subway-delay.

```{r}
#| label: tbl-code
#| tbl-cap: Sample of TTC Subway Delay Code
#| fig-pos: "H"
#| echo: false
#| message: false

@@

@@

```{r}
#| label: fig-total-delay
#| fig-cap: Visualization to show the TTC service that has the highest number of delays in year 2023
#| fig-pos: "H"
#| echo: false
#| message: false
#construct appropriate table for visualization
service_type <- c("bus", "streetcar", "subway")
num_delay <- c(nrow(bus_delay), nrow(streetcar_delay), nrow(subway_delay))

@@

```{r}
#| label: fig-weekday
#| fig-cap: Visualization to show the total delayed time during weekdays in 2023
#| fig-pos: "H"
#| echo: false
#| message: false

@@

```{r}
#| label: tbl-reason
#| tbl-cap: Top 5 Most Popular Reason for Delay
#| fig-pos: "H"
#| echo: false
#| message: false

@@

```{r}
#| label: tbl-lines
#| tbl-cap: Most delayed Lines for each TTC Service
#| fig-pos: "H"
#| echo: false
#| message: false
